---
output:
  html_document:
    keep_md: yes
---
Analysis of Fitbit Steps Data
-----------------------------------

To load the data,

```{r}
setwd("/Users/rossc/Dropbox/Engineering/RR1")
activity <- read.csv("activity.csv")
```

To find the total number of steps taken per day,

```{r}
by(activity$steps, activity$date, sum, na.rm = TRUE)
```

To plot the total number of steps taken per day in a histogram,

```{r}
total_steps <- by(activity$steps, activity$date, sum, na.rm = TRUE)
hist(total_steps, breaks = 15)
```

To take the mean and median of total steps per day,

```{r}
summary(total_steps)
```

To plot the mean number of steps taken over 60 days in each five minute interval,

```{r}
min <- levels(factor(activity$interval))
mean_steps <- by(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(min, mean_steps, type = "l")
```

To find the five minute interval with the maximum average number of steps, the first entry in the following will be the interval associated with the maximum:

```{r}
which.max(mean_steps)
```

To find the number of missing values (NA's) in the data,

```{r}
length(which(is.na(activity$steps)))
```

To create a new dataset with imputed values for all NA's called activity.complete,

```{r}
activity.complete <- activity
activity.complete$steps[is.na(activity.complete$steps)] = mean(activity.complete$steps, na.rm=TRUE)
```

To plot the total number of steps taken per day for activity.complete in a histogram,

```{r}
total_steps.complete <- by(activity.complete$steps, activity.complete$date, sum, na.rm = TRUE)
hist(total_steps.complete, breaks = 15)
```

To take the mean and median of total steps per day for activity.complete,

```{r}
summary(total_steps.complete)
```

Not surprisingly all the data becomes more squeezed towards the mean.

To look at whether the data was taken on a weekday or weekend we will add a new column with a True or False value to tell wether the observation was a weekday or not. Then we will change the values to Weekday and Weekend.

```{r}
activity.complete[, 4] <- !(weekdays(as.Date(activity.complete$date)) %in% c('Saturday', 'Sunday'))
activity.complete[activity.complete[, 4] == TRUE, 4] <- "Weekday"
activity.complete[activity.complete[, 4] == FALSE, 4] <- "Weekend"
names(activity.complete)[4] <- "weekdays"
```

To plot this data and compare the mean number of steps taken over the course of a weekday versus a weekend we use

```{r}
weekday.steps.mean <- by(activity.complete[activity.complete$weekdays == "Weekday" , 1], activity.complete[activity.complete$weekdays == "Weekday", 3], mean, na.rm = TRUE)
weekend.steps.mean <- by(activity.complete[activity.complete$weekdays == "Weekend" , 1], activity.complete[activity.complete$weekdays == "Weekend", 3], mean, na.rm = TRUE)
par(mfrow = c(2,1))
plot(weekday.steps.mean, type = "l", xlab = "Minutes", ylab = "Weekday Steps")
plot(weekend.steps.mean, type = "l", xlab = "Minutes", ylab = "Weekend Steps")
```